Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-03 latex/main.tex
Add brief Jackson explanation to Fig 3 caption

\begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{methodology.png}
    \caption{Schematic overview of sc2ts. (A) Procedure to build an ARG by incrementally updating it with samples sorted by collection dates. Sample 0 is used to initialize the ARG as the root. Next, Sample 1 is attached, mapping mutations onto the new branch to explain it. When Sample 4 is detected as a recombinant, a node representing the recombination event is inserted with two edges connecting to its parents, with Sample 4 itself attached to this node as its child. The resulting ARG contains two local trees, one on each side of the inferred breakpoint. (B) Copying paths for Sample 4 inferred under the LS model while allowing recombination ($k = 4$) or disallowing it ($k = 1000$).}
    \label{fig:methodology}
\end{figure}

@@

%%% TODO: Old version as placeholder. This figure is not yet finalised.
\begin{figure}[h]
    \centering
    \includegraphics[width=1\linewidth]{fig_recombinant_evidence.pdf}
    \caption{Evidence for recombination in the ARG. All plots show “robust” sc2ts recombination nodes, colored according
to their classification by rebar (recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table \ref{tab:pango_x_lineages} are labelled, point sizes reflect number of descendants;
Lower: recombination nodes classified by number of mutations averted by allowing recombination (i.e. $x$-$y$).
(B) Distribution of time to the MRCA of the parents of each robust recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns leading to XA, XBB, XBB.1, and the recombinant origin of XZ+XAC+XAD+XAE+XAP;
each recombination node is a mix of variants matching the left parent (green), the right parent (blue), and de-novo mutations (gold).}
    \label{fig:averted_mutations}
\end{figure}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{sc2ts_methodology}
    \caption{Schematic overview of sc2ts. (A) Procedure to build an ARG by incrementally updating it with samples sorted by collection dates. Sample 0 is used to initialize the ARG as the root. Next, Sample 1 is attached, mapping mutations onto the new branch to explain it. When Sample 4 is detected as a recombinant, a node representing the recombination event is inserted with two edges connecting to its parents, with Sample 4 itself attached to this node as its child. The resulting ARG contains two local trees, one on each side of the inferred breakpoint. (B) Copying paths for Sample 4 inferred under the LS model while allowing recombination ($k = 4$) or disallowing it ($k = 1000$).}
    \label{fig:methodology}
\end{figure}

@@

%%% TODO: Old version as placeholder. This figure is not yet finalised.
\begin{figure}[h]
    \centering
    \includegraphics[width=1\linewidth]{fig_recombinant_evidence.pdf}
    \caption{Evidence for recombination in the ARG. All plots show “robust” sc2ts recombination nodes, colored according
to their classification by rebar (recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table \ref{tab:pango_x_lineages} are labelled starting with `X' and
those identified by \cite{Jackson2021} (see Table \ref{tab:jackson}) are prefixed with a `J', e.g.
Jackson Group B (JGB) and Jackson Singleton recombinant 1 (JS1). 
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted by allowing recombination (i.e. $x$-$y$).
(B) Distribution of time to the MRCA of the parents of each robust recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns leading to XA, XBB, XBB.1, and the recombinant origin of XZ+XAC+XAD+XAE+XAP;
each recombination node is a mix of variants matching the left parent (green), the right parent (blue), and de-novo mutations (gold).}
    \label{fig:averted_mutations}
\end{figure}
